Weighted Model Counting Without Parameter Variables

### 2021-03-14 doc/paper3/paper.tex
Some edits before submission

\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{tikz}
\usepackage[binary-units]{siunitx}
\usepackage{booktabs}
\usepackage{capt-of}

@@

\begin{definition}[WMC]
  A \emph{WMC instance} is a tuple $(\phi, X_I, X_P, w)$, where $X_I$ is
  the set of \emph{indicator variables}, $X_P$ is the set of \emph{parameter
  variables} (with $X_I \cap X_P = \emptyset$), $\phi$ is a propositional formula
  in CNF over $X_I \cup X_P$, and $w\colon X_I \cup X_P \cup \{\neg x \mid x \in
  X_I \cup X_P\} \to \mathbb{R}$ is a \emph{weight function} such that $w(x) =
  w(\neg x) = 1$ for all $x \in X_I$. The \emph{answer} of the instance is
  $\sum_{Y \models \phi} \prod_{Y \models l} w(l)$.
\end{definition}

@@

\begin{definition}[Minimum-Cardinality WMC]
  A \emph{minimum-cardinality WMC} instance consists of the same tuple as a WMC
  instance, but its answer is defined to be $\sum_{Y \models \phi\text{, }|Y| =
    k} \prod_{Y \models l} w(l)$ (where $k = \min_{Y \models \phi} |Y|$) if
  $\phi$ is satisfiable, and zero otherwise.
\end{definition}

\begin{example}
  Let $\phi = (x \lor y) \land (\neg x \lor \neg y) \land (\neg x \lor p) \land
  (\neg y \lor q) \land x$, $X_I = \{ x, y \}$, $X_P = \{ p, q \}$, $w(p) =
  0.2$, $w(q) = 0.8$, and $w(\neg p) = w(\neg q) = 1$. Then $\phi$ has two
  models: $\{x, p\}$ and $\{ x, p, q \}$ with weights $0.2$ and $0.2 \times 0.8
  = 0.16$, respectively. The WMC answer is then $0.2 + 0.16 = 0.36$, and the
  minimum-cardinality WMC answer is $0.2$.
\end{example}

@@

WMC is a well-established technique for Bayesian network inference, particularly
effective on networks where most variables have only a few possible values
\cite{DBLP:conf/kr/Darwiche02}. Many ways of encoding a Bayesian network into a
WMC instance have been proposed. We will refer to them based on the initials of
the authors and the year of publication. Darwiche was the first to suggest the
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02} encoding that, in many ways, remains
the foundation behind most other encodings. He also introduced the distinction
between \emph{indicator} and \emph{parameter variables}; the former represent
variable-value pairs in the Bayesian network, while the latter are associated
with probabilities in the CPTs. The encoding \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05} is the only encoding that deviates from this
arrangement: for each variable in the Bayesian network, one indicator variable
acts simultaneously as a parameter variable. Chavira and Darwiche propose
\texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05} where they shift from WMC to
minimum-cardinality WMC because that allows the encoding to have fewer variables
and clauses. In particular, they propose a way to use the same parameter
variable to represent all probabilities in a CPT that are equal and keep only
clauses that `imply' parameter variables (i.e., omit clauses where a parameter
variable implies indicator variables). In their next encoding, \texttt{cd06}
\cite{DBLP:conf/sat/ChaviraD06}, the same authors optimise the aforementioned
implication clauses, choosing the smallest sufficient selection of indicator
variables. A decade later, Bart et al. present \texttt{bklm16}
\cite{DBLP:conf/ecai/BartKLM16} that improves upon \texttt{cd06} in two ways.
First, they optimise the number of indicator variables used per Bayesian network
variable from a linear to a logarithmic amount. Second, they introduce a scaling
factor that can `absorb' one probability per Bayesian network variable. However,
for this work, we choose to disable the latter improvement since this scaling
factor is often small enough to be indistinguishable from zero without the use
of arbitrary precision arithmetic. Indeed, even a small Bayesian network with
seven mutually independent binary variables, 0.1 and 0.9 probabilities each, is
already big enough for the scaling factor to be exactly equal to zero (as
produced by the \texttt{bklm16}
encoder\footnote{\url{http://www.cril.univ-artois.fr/kc/bn2cnf.html}}).
% TODO (on encodings and clauses): I realise the paper is already long but it
% might be useful to have a quick example
% TODO: If I am reading this correctly you are making a stark statement about
% the inefficacy. It might be good to rephrase this differently. You might want
% to say that there are many corner cases where there are no improvements. 
% TODO: Is there a reference for this example? If yes provide one. If no, maybe
% you could provide a link to the output, or alternatively, you could say
% something like the reader is free to check that such an example produces no
% improvements
% TODO: is this example the same as the previous one?

@@

\begin{definition}[PBP Instance] \label{def:new_wmc}
  A PBP instance is a tuple $(F, X, \omega)$, where $X$ is the set of variables,
  $F$ is a set of two-valued pseudo-Boolean functions $2^X \to \mathbb{R}$, and
  $\omega \in \mathbb{R}$ is the scaling factor.\footnote{
    Adding scaling factor $\omega$ to the definition allows us to remove clauses
    that consist entirely of a single parameter variable. The idea of extracting
    some of the structure of the WMC instance into an external multiplicative factor
    was loosely inspired by the \texttt{bklm16} encoding, where it is used
    to subsume the most commonly occurring probability of each CPT
    \cite{DBLP:conf/ecai/BartKLM16}.} Its \emph{answer} is $\omega \cdot
  \left(\exists_X\prod_{f \in F}f\right)(\emptyset)$.
\end{definition}

@@

\begin{proposition} \label{prop:equivalence}
  Let $(\phi, X_I, X_P, w)$ be a WMC instance. Then
  \begin{equation}
  \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
        x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
    1\right) \label{eq:new_wmc}
  \end{equation}
  is a PBP instance with the same answer.
\end{proposition}
% TODO: It’s a bit strange to say same answer, when you don’t precisely refer to
% this anywhere. One option is to say the above is the WMC instance and let
% Delta be the weighted model count. Then the new encoding Has a weighted model
% count of delta as well. Or something like that.
% But looking at the text below, I see that you use this phrasing everywhere,
% and I admit my formulation is awkward. So maybe the best compromise here is
% just simply add a footnote explaining what exactly you mean when you say same
% answer
% TODO: If it is the case that you cannot include the proofs in line, and indeed
% it might be better for readability to have an appendix with all the proofs, it
% might make sense to provide the type of argument you are using in the proofs
% so that readers know what to expect when they look into it
\begin{proof}
  Let $f = \prod_{c \in \phi} [c]_0^1$, and $g = \prod_{x \in X_I \cup X_P}
  [x]_{w(\neg x)}^{w(x)}$. Then the WMC answer is
  \eqref{eq:new_wmc} is
  \[
    (\exists_{X_I \cup X_P} fg)(\emptyset) = \sum_{Y \subseteq X_I \cup X_P}
    (fg)(Y) = \sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y)
  \]
  by \cref{lemma:sum}. Note that
  \[
    f(Y) =
    \begin{cases}
      1 & \text{if } Y \models \phi, \\
      0 & \text{otherwise},
    \end{cases}
    \quad
    \text{and}
    \quad
    g(Y) = \prod_{Y \models l} w(l),
  \]
  which means that $\sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y) = \sum_{Y \models
    \phi} \prod_{Y \models l} w(l)$ as required.
\end{proof}

\begin{theorem}[Correctness for WMC] \label{thm:correctness}
  \Cref{alg:transformation}, when given a WMC instance $(\phi, X_I, X_P,
  w)$, returns a PBP instance with the same answer, provided \emph{either} of
  the two conditions is satisfied:
  \begin{enumerate}
  \item for all $p \in X_P$, there is a non-empty family of literals
    $(l_i)_{i=1}^n$ such that \label{cond:d02}
    \begin{enumerate}
    \item $w(\neg p) = 1$,
    \item $l_i \in X_I$ or $\neg l_i \in X_I$ for all $i = 1, \dots, n$,
      \label{condition:equivalence1}
    \item and $\{ c \in \phi \mid p \in c \text{ or } \neg p \in c \} =
      \left\{p \lor \bigvee_{i=1}^n \neg l_i \right\} \cup \{ l_i \lor \neg p
      \mid i = 1, \dots, n \}$; \label{condition:equivalence2}
    \end{enumerate}
  \item or for all $p \in X_P$, \label{cond:bklm16}
    \begin{enumerate}
    \item $w(p) + w(\neg p) = 1$,
    \item for any clause $c \in \phi$, $|c \cap X_P| \le 1$, \label{cond:2b2}
    \item there is no clause $c \in \phi$ such that $\neg p \in
      c$, \label{cond:2b3}
    \item if $\{ p \} \in \phi$, then there is no clause $c \in \phi$ such
      that $c \ne \{ p \}$ and $p \in c$, \label{cond:just_parameter}
    \item and for any $c, d \in \phi$ such that $c \ne d$, $p \in c$ and $p
      \in d$, $\bigwedge_{l \in c \setminus \{ p \}} \neg l \land \bigwedge_{l
        \in d \setminus \{ p \}} \neg l$ is false. \label{cond:disjoint}
    \end{enumerate}
  \end{enumerate}
\end{theorem}

@@

\begin{proof}
  By \cref{prop:equivalence},
  \begin{equation}
    \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
          x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
      1\right) \label{eq:new_wmc2}
  \end{equation}
  is a PBP instance with the same answer as the given WMC instance. By
  \cref{def:new_wmc}, its answer is $\left(\exists_{X_I \cup X_P} \left(\prod_{c
        \in \phi} [c]_0^1 \right) \prod_{x \in X_I \cup X_P} [x]_{w(\neg
      x)}^{w(x)} \right)(\emptyset)$. Since both \cref{cond:d02,cond:bklm16}
  ensure that each clause in $\phi$ has at most one parameter variable, we can
  partition $\phi$ into $\phi_* \coloneqq \{c \in \phi \mid \mathtt{Vars}(c)
  \cap X_P = \emptyset \}$ and $\phi_p \coloneqq \{ c \in \phi \mid
  \mathtt{Vars}(c) \cap X_P = \{ p \} \}$ for all $p \in X_P$. We can then use
  \cref{thm:early} to reorder the answer into $\left(\exists_{X_I} \left(
      \prod_{x \in X_I} [x]_{w(\neg x)}^{w(x)} \right) \left( \prod_{c \in
        \phi_*} [c]_0^1 \right) \prod_{p \in X_P} \exists_p [p]_{w(\neg
      p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 \right)(\emptyset)$.

@@

  If $w(p) \ne 1$, and $\{ p \} \in \phi_p$, then, by
  \cref{cond:just_parameter}, $\phi_p = \{ \{ p \} \}$, and
  \cref{alg:transformation} produces $F_p = \emptyset$ and $\omega = w(p)$, and
  so
  \[
    \exists_p [p]_{w(\neg p)}^{w(p)} [p]_0^1 = \exists_p [p]^{w(p)}_0 = w(p) =
    \omega \prod_{f \in F_p} f.
  \]
  The only remaining case is when $w(p) \ne 1$ and $\{ p \} \not \in \phi_p$.
  Then $\omega = 1$, and $F_p = \left\{ \left[\bigwedge_{l \in c \setminus \{ p
        \}} \neg l\right]_1^{w(p)} \;\middle|\; c \in \phi_p \right\}$, so we
  need to show that
  \[
    \prod_{c \in \phi_p} \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
      l\right]_1^{w(p)} = \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p}
    [c]_0^1.
  \]
  We can rearrange the right-hand side as
  % \begin{align*}
  %   \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[\bigwedge_{c \in \phi_p} c\right]_0^1 = \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
  %                                                              &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 = w(p) + \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^{1-w(p)} \\
  %                                                              &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \neg \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \neg(c \setminus \{ p \}) \right]_1^{w(p)} \\
  %                                                              &= \left[ \bigvee_{c \in \phi_p} \neg \bigvee_{l \in c \setminus \{ p \}} l \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}.
  % \end{align*}
  \begin{align*}
    \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}.
  \end{align*}
   By \cref{cond:disjoint}, $\bigwedge_{l \in c \setminus \{ p \}} \neg l$ can
   be true for at most one $c \in \phi_p$, and so
   \[
     \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l
     \right]_1^{w(p)} = \prod_{c \in \phi_p} \left[ \bigwedge_{l \in c \setminus
         \{ p \}} \neg l \right]_1^{w(p)}
   \]
   which is exactly what we needed to show. This ends the proof that the
   first loop of \cref{alg:transformation} preserves the answer under both
   \cref{cond:d02} and \cref{cond:bklm16}. Finally, the loop on
   \crefrange{line:foreach2start}{line:foreach2end} of \cref{alg:transformation}
   replaces $[v]_1^p[\neg v]_1^q$ with $[v]_q^p$ (for some $v \in X_I$ and $p, q
   \in \mathbb{R}$), but, of course, $[v]_1^p[\neg v]_1^q = [v]_1^p[v]_q^1 =
   [v]_q^p$, i.e., the answer is unchanged.
\end{proof}

\begin{theorem}[Minimum-Cardinality Correctness] \label{thm:mccorrectness}
  Let $(\phi, X_I, X_P, w)$ be a minimum-cardinality WMC instance that satisfies
  \cref{cond:2b2,cond:2b3,cond:just_parameter,cond:disjoint} of
  \cref{thm:correctness} as well as the following:
  \begin{enumerate}
  \item for all parameter variables $p \in X_P$, $w(\neg p) = 1$.
  \item all models of $\{ c \in \phi \mid c \cap X_P = \emptyset \}$ (as subsets
    of $X_I$) have the same cardinality; \label{cond:22}
  \item $\min_{Z \subseteq X_P} |Z|$ such that $Y \cup Z \models \phi$ is the
    same for all $Y \models \{ c \in \phi \mid c \cap X_P = \emptyset
    \}$. \label{cond:23}
  \end{enumerate}
  Then \cref{alg:transformation}, when applied to $(\phi, X_I, X_P, w)$, outputs
  a PBP instance with the same answer.
\end{theorem}

@@

For experimental data, we use Bayesian networks available with \textsf{Ace} and
\textsf{Cachet}. We split them into the following groups:
\begin{itemize}
\item DQMR (390 instances) and
\item Grid networks (450 instances) as described by Sang et al.
  \cite{DBLP:conf/aaai/SangBK05};
\item Mastermind (144 instances) and
\item Random Blocks (256 instances) by Chavira et al.
  \cite{DBLP:journals/ijar/ChaviraDJ06};
\item other binary Bayesian networks (50 instances) including Plan Recognition
  \cite{DBLP:conf/aaai/SangBK05}, Friends and Smokers, Students and Professors
  \cite{DBLP:journals/ijar/ChaviraDJ06}, and \texttt{tcc4f};
\item non-binary classic networks (176 instances): \texttt{alarm},
  \texttt{diabetes}, \texttt{hailfinder}, \texttt{mildew},
  \texttt{munin1}--\texttt{4}, \texttt{pathfinder}, \texttt{pigs}, and
  \texttt{water}.
\end{itemize}

>>>>

\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{tikz}
\usepackage[binary-units]{siunitx}
\usepackage{booktabs}
\usepackage{capt-of}
\usepackage[inline]{enumitem}

@@

\begin{definition}[WMC] \label{def:wmc}
  A \emph{WMC instance} is a tuple $(\phi, X_I, X_P, w)$, where $X_I$ is
  the set of \emph{indicator variables}, $X_P$ is the set of \emph{parameter
  variables} (with $X_I \cap X_P = \emptyset$), $\phi$ is a propositional formula
  in CNF over $X_I \cup X_P$, and $w\colon X_I \cup X_P \cup \{\neg x \mid x \in
  X_I \cup X_P\} \to \mathbb{R}$ is a \emph{weight function} such that $w(x) =
  w(\neg x) = 1$ for all $x \in X_I$. The \emph{answer} of the instance is
  $\sum_{Y \models \phi} \prod_{Y \models l} w(l)$.
\end{definition}

@@

\begin{definition}[Minimum-Cardinality WMC] \label{def:mcwmc}
  A \emph{minimum-cardinality WMC} instance consists of the same tuple as a WMC
  instance, but its \emph{answer} is defined to be $\sum_{Y \models \phi\text{,
    }|Y| = k} \prod_{Y \models l} w(l)$ (where $k = \min_{Y \models \phi} |Y|$)
  if $\phi$ is satisfiable, and zero otherwise.
\end{definition}

\begin{example} \label{example:1}
  Let $\phi = (x \lor y) \land (\neg x \lor \neg y) \land (\neg x \lor p) \land
  (\neg y \lor q) \land x$, $X_I = \{ x, y \}$, $X_P = \{ p, q \}$, $w(p) =
  0.2$, $w(q) = 0.8$, and $w(\neg p) = w(\neg q) = 1$. Then $\phi$ has two
  models: $\{x, p\}$ and $\{ x, p, q \}$ with weights $0.2$ and $0.2 \times 0.8
  = 0.16$, respectively. The WMC answer is then $0.2 + 0.16 = 0.36$, and the
  minimum-cardinality WMC answer is $0.2$.
\end{example}

@@

WMC is a well-established technique for Bayesian network inference, particularly
effective on networks where most variables have only a few possible values
\cite{DBLP:conf/kr/Darwiche02}. Many ways of encoding a Bayesian network into a
WMC instance have been proposed. We will refer to them based on the initials of
the authors and the year of publication. Darwiche was the first to suggest the
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02} encoding that, in many ways, remains
the foundation behind most other encodings. He also introduced the distinction
between \emph{indicator} and \emph{parameter variables}; the former represent
variable-value pairs in the Bayesian network, while the latter are associated
with probabilities in the CPTs. The encoding \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05} is the only encoding that deviates from this
arrangement: for each variable in the Bayesian network, one indicator variable
acts simultaneously as a parameter variable. Chavira and Darwiche propose
\texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05} where they shift from WMC to
minimum-cardinality WMC because that allows the encoding to have fewer variables
and clauses. In particular, they propose a way to use the same parameter
variable to represent all probabilities in a CPT that are equal and keep only
clauses that `imply' parameter variables (i.e., omit clauses where a parameter
variable implies indicator variables).\footnote{\Cref{example:2} demonstrates
  what we mean by implication clauses.} In their next encoding, \texttt{cd06}
\cite{DBLP:conf/sat/ChaviraD06}, the same authors optimise the aforementioned
implication clauses, choosing the smallest sufficient selection of indicator
variables. A decade later, Bart et al. present \texttt{bklm16}
\cite{DBLP:conf/ecai/BartKLM16} that improves upon \texttt{cd06} in two ways.
First, they optimise the number of indicator variables used per Bayesian network
variable from a linear to a logarithmic amount. Second, they introduce a scaling
factor that can `absorb' one probability per Bayesian network variable. However,
for this work, we choose to disable the latter improvement since this scaling
factor is often small enough to be indistinguishable from zero without the use
of arbitrary precision arithmetic, making it completely unusable on realistic
instances. Indeed, the reader is free to check that even a small Bayesian
network with seven mutually independent binary variables, 0.1 and 0.9
probabilities each, is already big enough for the scaling factor to be exactly
equal to zero (as produced by the \texttt{bklm16}
encoder\footnote{\url{http://www.cril.univ-artois.fr/kc/bn2cnf.html}}). We
suspect that this issue was not identified during the original set of
experiments because they never looked at numerical answers.

\begin{example} \label{example:2}
  Let $\mathcal{B}$ be a Bayesian network with one variable $X$ which has two
  values $x_1$ and $x_2$ with probabilities $\Pr(X = x_1) = 0.2$ and $\Pr(X =
  x_2) = 0.8$. Let $x, y$ be indicator variables, and $p, q$ be parameter
  variables. Then \cref{example:1} is both the \texttt{cd05} and the
  \texttt{cd06} encoding of $\mathcal{B}$. The \texttt{bklm16} encoding is $(x
  \Rightarrow p) \land (\neg x \Rightarrow q) \land x$ with $w(p) = w(\neg q) =
  0.2$, and $w(\neg p) = w(q) = 0.8$. And the \texttt{d02} encoding is $(\neg x
  \Rightarrow p) \land (p \Rightarrow \neg x) \land (x \Rightarrow q) \land (q
  \Rightarrow x) \land \neg x$ with $w(p) = 0.2$, $w(q) = 0.8$, and $w(\neg p) =
  w(\neg q) = 1$. Note how all other encodings have fewer clauses than
  \texttt{d02}. While \texttt{cd05} and \texttt{cd06} require
  minimum-cardinality WMC to make this work, \texttt{bklm16} achieves the same
  thing by adjusting weights.
\end{example}

@@

\begin{definition}[PBP Instance] \label{def:pbp}
  A PBP instance is a tuple $(F, X, \omega)$, where $X$ is the set of variables,
  $F$ is a set of two-valued pseudo-Boolean functions $2^X \to \mathbb{R}$, and
  $\omega \in \mathbb{R}$ is the scaling factor.\footnote{
    Adding scaling factor $\omega$ to the definition allows us to remove clauses
    that consist entirely of a single parameter variable. The idea of extracting
    some of the structure of the WMC instance into an external multiplicative factor
    was loosely inspired by the \texttt{bklm16} encoding, where it is used
    to subsume the most commonly occurring probability of each CPT
    \cite{DBLP:conf/ecai/BartKLM16}.} Its \emph{answer} is $\omega \cdot
  \left(\exists_X\prod_{f \in F}f\right)(\emptyset)$.
\end{definition}

@@

\begin{proposition} \label{prop:equivalence}
  Let $(\phi, X_I, X_P, w)$ be a WMC instance. Then
  \begin{equation}
  \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
        x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
    1\right) \label{eq:new_wmc}
  \end{equation}
  is a PBP instance with the same answer (as defined in \cref{def:wmc,def:pbp}).
\end{proposition}
% It’s a bit strange to say same answer, when you don’t precisely refer to
% this anywhere. One option is to say the above is the WMC instance and let
% Delta be the weighted model count. Then the new encoding Has a weighted model
% count of delta as well. Or something like that.
% But looking at the text below, I see that you use this phrasing everywhere,
% and I admit my formulation is awkward. So maybe the best compromise here is
% just simply add a footnote explaining what exactly you mean when you say same
% answer
\begin{proof}
  Let $f = \prod_{c \in \phi} [c]_0^1$, and $g = \prod_{x \in X_I \cup X_P}
  [x]_{w(\neg x)}^{w(x)}$. Then the WMC answer is
  \eqref{eq:new_wmc} is
  \[
    (\exists_{X_I \cup X_P} fg)(\emptyset) = \sum_{Y \subseteq X_I \cup X_P}
    (fg)(Y) = \sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y)
  \]
  by \cref{lemma:sum}. Note that
  \[
    f(Y) =
    \begin{cases}
      1 & \text{if } Y \models \phi, \\
      0 & \text{otherwise},
    \end{cases}
    \quad
    \text{and}
    \quad
    g(Y) = \prod_{Y \models l} w(l),
  \]
  which means that $\sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y) = \sum_{Y \models
    \phi} \prod_{Y \models l} w(l)$ as required.
\end{proof}

\begin{theorem}[Correctness for WMC] \label{thm:correctness}
  \Cref{alg:transformation}, when given a WMC instance $(\phi, X_I, X_P,
  w)$, returns a PBP instance with the same answer (as defined in
  \cref{def:wmc,def:pbp}), provided \emph{either} of the two conditions is
  satisfied:
  \begin{enumerate}
  \item for all $p \in X_P$, there is a non-empty family of literals
    $(l_i)_{i=1}^n$ such that \label{cond:d02}
    \begin{enumerate}
    \item $w(\neg p) = 1$,
    \item $l_i \in X_I$ or $\neg l_i \in X_I$ for all $i = 1, \dots, n$,
      \label{condition:equivalence1}
    \item and $\{ c \in \phi \mid p \in c \text{ or } \neg p \in c \} =
      \left\{p \lor \bigvee_{i=1}^n \neg l_i \right\} \cup \{ l_i \lor \neg p
      \mid i = 1, \dots, n \}$; \label{condition:equivalence2}
    \end{enumerate}
  \item or for all $p \in X_P$, \label{cond:bklm16}
    \begin{enumerate}
    \item $w(p) + w(\neg p) = 1$,
    \item for any clause $c \in \phi$, $|c \cap X_P| \le 1$, \label{cond:2b2}
    \item there is no clause $c \in \phi$ such that $\neg p \in
      c$, \label{cond:2b3}
    \item if $\{ p \} \in \phi$, then there is no clause $c \in \phi$ such
      that $c \ne \{ p \}$ and $p \in c$, \label{cond:just_parameter}
    \item and for any $c, d \in \phi$ such that $c \ne d$, $p \in c$ and $p
      \in d$, $\bigwedge_{l \in c \setminus \{ p \}} \neg l \land \bigwedge_{l
        \in d \setminus \{ p \}} \neg l$ is false. \label{cond:disjoint}
    \end{enumerate}
  \end{enumerate}
\end{theorem}

@@

\begin{proof}
  By \cref{prop:equivalence},
  \begin{equation}
    \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
          x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
      1\right) \label{eq:new_wmc2}
  \end{equation}
  is a PBP instance with the same answer as the given WMC instance. By
  \cref{def:pbp}, its answer is $\left(\exists_{X_I \cup X_P} \left(\prod_{c
        \in \phi} [c]_0^1 \right) \prod_{x \in X_I \cup X_P} [x]_{w(\neg
      x)}^{w(x)} \right)(\emptyset)$. Since both \cref{cond:d02,cond:bklm16}
  ensure that each clause in $\phi$ has at most one parameter variable, we can
  partition $\phi$ into $\phi_* \coloneqq \{c \in \phi \mid \mathtt{Vars}(c)
  \cap X_P = \emptyset \}$ and $\phi_p \coloneqq \{ c \in \phi \mid
  \mathtt{Vars}(c) \cap X_P = \{ p \} \}$ for all $p \in X_P$. We can then use
  \cref{thm:early} to reorder the answer into $\left(\exists_{X_I} \left(
      \prod_{x \in X_I} [x]_{w(\neg x)}^{w(x)} \right) \left( \prod_{c \in
        \phi_*} [c]_0^1 \right) \prod_{p \in X_P} \exists_p [p]_{w(\neg
      p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 \right)(\emptyset)$.

@@

  If $w(p) \ne 1$, and $\{ p \} \in \phi_p$, then, by
  \cref{cond:just_parameter}, $\phi_p = \{ \{ p \} \}$, and
  \cref{alg:transformation} produces $F_p = \emptyset$ and $\omega = w(p)$, and
  so
  \[
    \exists_p [p]_{w(\neg p)}^{w(p)} [p]_0^1 = \exists_p [p]^{w(p)}_0 = w(p) =
    \omega \prod_{f \in F_p} f.
  \]
  The only remaining case is when $w(p) \ne 1$ and $\{ p \} \not \in \phi_p$.
  Then $\omega = 1$, and $F_p = \left\{ \left[\bigwedge_{l \in c \setminus \{ p
        \}} \neg l\right]_1^{w(p)} \;\middle|\; c \in \phi_p \right\}$, so we
  need to show that $\prod_{c \in \phi_p} \left[\bigwedge_{l \in c \setminus \{
      p \}} \neg l\right]_1^{w(p)} = \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in
    \phi_p} [c]_0^1$. We can rearrange the right-hand side as
  % \begin{align*}
  %   \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[\bigwedge_{c \in \phi_p} c\right]_0^1 = \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
  %                                                              &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 = w(p) + \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^{1-w(p)} \\
  %                                                              &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \neg \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \neg(c \setminus \{ p \}) \right]_1^{w(p)} \\
  %                                                              &= \left[ \bigvee_{c \in \phi_p} \neg \bigvee_{l \in c \setminus \{ p \}} l \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}.
  % \end{align*}
  \begin{align*}
    \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}.
  \end{align*}
   By \cref{cond:disjoint}, $\bigwedge_{l \in c \setminus \{ p \}} \neg l$ can
   be true for at most one $c \in \phi_p$, and so $\left[ \bigvee_{c \in \phi_p}
     \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)} = \prod_{c
     \in \phi_p} \left[ \bigwedge_{l \in c \setminus \{ p \}} \neg l
   \right]_1^{w(p)}$ which is exactly what we needed to show. This ends the
   proof that the first loop of \cref{alg:transformation} preserves the answer
   under both \cref{cond:d02} and \cref{cond:bklm16}. Finally, the loop on
   \crefrange{line:foreach2start}{line:foreach2end} of \cref{alg:transformation}
   replaces $[v]_1^p[\neg v]_1^q$ with $[v]_q^p$ (for some $v \in X_I$ and $p, q
   \in \mathbb{R}$), but, of course, $[v]_1^p[\neg v]_1^q = [v]_1^p[v]_q^1 =
   [v]_q^p$, i.e., the answer is unchanged.
\end{proof}

\begin{theorem}[Minimum-Cardinality Correctness] \label{thm:mccorrectness}
  Let $(\phi, X_I, X_P, w)$ be a minimum-cardinality WMC instance that satisfies
  \cref{cond:2b2,cond:2b3,cond:just_parameter,cond:disjoint} of
  \cref{thm:correctness} as well as the following:
  \begin{enumerate}
  \item for all parameter variables $p \in X_P$, $w(\neg p) = 1$.
  \item all models of $\{ c \in \phi \mid c \cap X_P = \emptyset \}$ (as subsets
    of $X_I$) have the same cardinality; \label{cond:22}
  \item $\min_{Z \subseteq X_P} |Z|$ such that $Y \cup Z \models \phi$ is the
    same for all $Y \models \{ c \in \phi \mid c \cap X_P = \emptyset
    \}$. \label{cond:23}
  \end{enumerate}
  Then \cref{alg:transformation}, when applied to $(\phi, X_I, X_P, w)$, outputs
  a PBP instance with the same answer (as defined in \cref{def:mcwmc,def:pbp}).
\end{theorem}

@@

For experimental data, we use Bayesian networks available with \textsf{Ace} and
\textsf{Cachet}. We split them into the following groups:
\begin{itemize*}
\item DQMR (390 instances) and
\item Grid networks (450 instances) as described by Sang et al.
  \cite{DBLP:conf/aaai/SangBK05};
\item Mastermind (144 instances) and
\item Random Blocks (256 instances) by Chavira et al.
  \cite{DBLP:journals/ijar/ChaviraDJ06};
\item other binary Bayesian networks (50 instances) including Plan Recognition
  \cite{DBLP:conf/aaai/SangBK05}, Friends and Smokers, Students and Professors
  \cite{DBLP:journals/ijar/ChaviraDJ06}, and \texttt{tcc4f};
\item non-binary classic networks (176 instances): \texttt{alarm},
  \texttt{diabetes}, \texttt{hailfinder}, \texttt{mildew},
  \texttt{munin1}--\texttt{4}, \texttt{pathfinder}, \texttt{pigs}, and
  \texttt{water}.
\end{itemize*}
